Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development

### 2024-03-24 paper.qmd

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Toole"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach to the development of strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the methods used to generate the evidence presented in the tool, including disaggregation of origin-destination data, additional trip purposes, routing, scenarios of cycling potential, and the user interface. The CRUSE Tool provides estimates of cycling currently and under four scenarios of cycling potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. The tool was commissioned by Transport Infrastructure Ireland and is publicly available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modelling networks at high levels of geographic resolution; simulating multiple trip purposes (social, shopping and personal utility trips plus recreational and cycle touring trips), in addition to using official data on trips to work and education locations; provision of estimates of 'quietness' (a proxy for cyclability) at the route segment level; and its provision of three network types: 'Fastest', 'Balanced' and 'Quietest' to help plan both arterial and residential cycle networks. Workshops and unstructured interviews of stakeholders show that the tool is already being used in practice, to inform urban, inter-urban and rural network development. The approach and methods for strategic cycle network development tools outlined in this paper can be applied and built-on in other places to support the generation of evidence needed to support ambitious cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---

>>>>

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Toole"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the methods used to generate the evidence presented in the tool, including disaggregation of origin-destination data, additional trip purposes, routing, scenarios of cycling potential, and the user interface. The CRUSE Tool provides estimates of cycling currently and under four scenarios of cycling potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. The tool was commissioned by Transport Infrastructure Ireland and is publicly available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modelling networks at high levels of geographic resolution; simulating multiple trip purposes (social, shopping and personal utility trips plus recreational and cycle touring trips), in addition to using official data on trips to work and education locations; provision of estimates of 'quietness' (a proxy for cyclability) at the route segment level; and its provision of three network types: 'Fastest', 'Balanced' and 'Quietest' to help plan both arterial and residential cycle networks. Workshops and unstructured interviews of stakeholders show that the tool is already being used in practice, to inform urban, inter-urban and rural network development. The approach and methods for strategic cycle network development tools outlined in this paper can be applied and built-on in other places to support the generation of evidence needed to support ambitious cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---
